---
title: "Imputation Dataset"
output: html_document
---

```{r}
library(mice)
library(tidyverse)
```

```{r}
full_data <- read.csv("full_data.csv")
```

Start with imputing the data. We know from earlier analysis that there is a large proportion of the variables in the dataset missing a lof of observations. If we do not impute the data, then it makes it very difficult/ impossible to perform analyses on the variables in the dataset. 

Look at the missing data:
```{r}
aggr_plot <- aggr(full_data, col=c('darkblue','darkred'), numbers=TRUE, sortVars=TRUE, labels=names(data), cex.axis=.7, gap=3, ylab=c("Histogram of missing data","Pattern"))


```
From the plots above we can see a large number of the variables are missing over 80% of their observations and many more mising well above 50% of their observations.


```{r}
#impute data
#we need to seperate the variables out that we don't want imputed on like identification variables
imputed_data<- full_data %>%
  dplyr::select(-c("Code", "TeacherID", "School_Type", "Level", "KnowledgePercent_ScoreChange", "KnowledgePercent"))
not_imputed<- full_data %>%
  dplyr::select(c("Code", "TeacherID", "School_Type", "Level", "KnowledgePercent_ScoreChange", "KnowledgePercent"))
```

Use the mice package to impute the data
```{r}
set.seed(234)
# step1: replace the missing value with the values of the first imputation to create a complete dataset 
newdata <- mice(imputed_data, m=5, maxit=10, method="pmm", printFlag = FALSE) # there are more methods of multiple imputation you can try, not only pmm, to see more detail, check: https://www.rdocumentation.org/packages/mice/versions/3.13.0/topics/mice

newdata2 <- complete(newdata, 1, inc = FALSE)
newdata2 <- cbind(newdata2, not_imputed)
#should have 2351 logged events (or NAs imputed)
```


```{r}
#remove unecessary columns in newdata2
newdata2 %<>%
  dplyr::select(-c(`X`,`GenScoreV1`, `GenScoreV2`,`school_type`,`school`,`grade_oth`,`kit`,`preweight`,`presmoke`,`preasthma`,`preengage`,`post`,`poststress`,`postjobstress`,`program_pd1`,`program_pd2`,`program_pd3`,`program_pd4`))
#should now have 77 variables
```

```{r}
write.csv(newdata2, "newdata2.csv")
```
