---
title: "Teacher_Cleaning"
author: "Lauren A. Kohls and Sarah Little"
date: "4/26/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE,warning=FALSE, message=FALSE, error=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message=FALSE)
```


```{r message=FALSE, warning=FALSE, include=FALSE}
# Load Libraries
library(foreign)
library(tidyverse)
library(ggplot2)
library(magrittr)
library(dplyr)
```

##Section1: Initial Cleaning


###Step 1: read in teacher spss file

```{r error=FALSE, message=FALSE, include=FALSE}

#read in Teacher data SPSS file 
teacher_data <- data.frame(read.spss('Teacher Data_March_2020_Adrian .sav'))
teacher_data2 <- teacher_data #create duplicate dataset
#rm(teacher_data)


```

### Step 2: Remove NAs 
We can see that there are a large amount of schools missing information. Out of the 206 observations, 132 school names are missing. Hopefully when we merge the teacher data with the student data, the missing school names should be accounted for. 

We only see 3 rows with missing data for all observations at the same time, so we look to remove those rows. After removing the 3 entrees, we see the dataset decreased from 206 observations to 203 observations, the exact result we wanted.

```{r include=FALSE}
#search for NAs in the code, school type and Y1_School variables because we want to eventually merge this data set with the student data 
data.frame(Num_NA = c(teacher_data2 %>% filter(is.na(Code) | Code == "NA") %>% nrow(),
                      teacher_data2 %>% filter(is.na(School_Type) | School_Type == "NA") %>% nrow(),teacher_data2 %>% filter(is.na(Y1_School) | Y1_School == "NA") %>% nrow()))

```
```{r include=FALSE}
#remove teacher_data
rm(teacher_data)
```

```{r include=FALSE}

#remove NAs from the duplicate dataset teacher_data2
teacher_clean <- teacher_data2 %>% 
  filter(!is.na(Code) & Code != "" &
                                !is.na(School_Type) & School_Type != "")
teacher_clean %>% 
  nrow() # Count the remaining entries in the data set

```

### Step 3: Remove unused Variables: 
```{r include=FALSE}
#now look to remove unused variables (this is very tedious, might be better in the future to just select the variables we definitly want to ise )

#also remove variables that are duplicates (ie there is a column that combines feet and inches, so repetative to have both)

#also remove columns that have no entries (ie PostPD_8Other)


teacher_clean <- teacher_clean %>% dplyr::select(-contains("y4")) #there is no year 4 data currently, but in the future this line of code will need to be removed

teacher_clean <- teacher_clean %>% dplyr::select(-contains("Y1_PostRace_2"))
teacher_clean <- teacher_clean %>% dplyr::select(-contains("Y1_PostRace_3"))
teacher_clean <- teacher_clean %>% dplyr::select(-contains("Y1_PostPD_8Other"))
teacher_clean <- teacher_clean %>% dplyr::select(-contains("y2_PostRace_2"))
teacher_clean <- teacher_clean %>% dplyr::select(-contains("y2_PostRace_3"))
teacher_clean <- teacher_clean %>% dplyr::select(-contains("y2_PostPD_8Other"))

#we now have only observations for years 1, 2, and 3 and 503 variables, down from the 626 original variables
rm(teacher_data2)
```



##Section 2: Drop Columns

```{r include=FALSE}
drops <- c("Y1_PreRace_Other","Y1_PreRace_2", "Y1_PreRace_3", "Y1_PreHispanic","Y1_PreHealthNumShift","Y1_PrePhyHealth","Y1_PreStressInv", "Y1_PreStressInvNumShift", "Y1_PrePhyActive", "Y1_PreSit", "Y1_PreComputer", "Y1_PreVig_Intensity", "Y1_PreMod_Intensity", "Y1_PreFitBit", "Y1_PreSoda", "Y1_PreFruit", "Y1_PreVegatables", "Y1_PreWater", "Y1_PreBreakfast", "Y1_PreDeliver_NutEd", "Y1_PreOpen_NutEd", "Y1_PreOther_NutEd","Y1_PreOther_NutEd2", "Y1_PreStudentHealthNumShift","Y1_PreTeacherRespNumShift","Y1_PreIntegrateNumShift", "Y1_PreParticipateNumeShift", "Y1_PreImpactNumShift", "Y1_PreAchievementNumShift", "Y1_PrePromoteNumShift", "Y1_PreEngageNumShift", "Y1_PreValueNumShift", "Y1_PreMotivateNumShift", "Y1_PreAssistNumShift", "Y1_PreImproveNumShift", "Y1_PreThinkNumShift", "Y1_PrePD_1Physical", "Y1_PrePD_2Stress", "Y1_PrePD_3Flu", "Y1_PrePD_4Nutrition", "Y1_PrePD_5Meditatio", "Y1_PrePD_6Smoking", "Y1_PrePD_7Ergo", "Y1_PrePD_8Other", "Y1_PrePD_Other", "Y1_PostRace_Other", "Y1_PostHealthNumShift", "Y1_PostPhyHealth", "Y1_PostMentalHealth", "Y1_PostStressInvNumShift", "Y1_PostSmoke", "Y1_PostTobProduct", "Y1_PostDiabetes", "Y1_PostHypertention", "Y1_PostAsthma", "Y1_PostPhyActive", "Y1_PostSit", "Y1_PostComputer","Y1_PostVig_Intensity", "Y1_PostMod_Intensity", "Y1_PostFitBit", "Y1_PostSoda", "Y1_PostFruit", "Y1_PostVegatables", "Y1_PostWater", "Y1_PostBreakfast", "Y1_PostDeliver_NutEd", "Y1_PostOpen_NutEd", "Y1_PostOther_NutEd","Y1_PostOther_NutEd2","Y1_PostStudentHealthNumShift","Y1_PostTeacherRespNumShift","Y1_PostIntegrateNumShift", "Y1_PostParticipateNumeShift", "Y1_PostImpactNumShift", "Y1_PostAchievementNumShift", "Y1_PostPromoteNumShift", "Y1_PostEngageNumShift", "Y1_PostValueNumShift","Y1_PostMotivateNumShift", "Y1_PostAssistNumShift", "Y1_PostImproveNumShift", "Y1_PostThinkNumShift","Y1_PostPD_1Physical", "Y1_PostPD_2Stress", "Y1_PostPD_3Flu", "Y1_PostPD_4Nutrition", "Y1_PostPD_5Meditatio", "Y1_PostPD_6Smoking", "Y1_PostPD_7Ergo", "Y1_PostPD_8Other", "Y1_PostPD_Other", "Y1_Program_NotChange", "Y1_Program_Change", "y2_prerace_other", "y2_prerace_2", "y2_prerace_3", "y2_prehispanic", "y2_PreHealthNumShift", "y2_prephyhealth", "y2_prementhealth", "y2_prestress", "y2_prestressinv", "y2_PreStressInvNumShift", "y2_prejobstress", "y2_presmoke", "y2_pretobproduct", "y2_presleep", "y2_prediabetes", "y2_prehypertension", "y2_preasthma", "y2_prephyactive", "y2_presit", "y2_precomputer", "y2_previgintensity", "y2_premod_intensity", "y2_prefitbit", "y2_presoda", "y2_prefruit", "y2_prevegtables", "y2_prewater", "y2_prebreakfast", "y2_predeliver_nuted", "y2_preopen_nuted", "y2_preother_nuted", "y2_preother_nuted2", "y2_prehealthed", "y2_prestudenthealth", "y2_preteacherresp", "y2_preintegrate", "y2_preparticipate", "y2_preimpact", "y2_preachievement", "y2_prepromote", "y2_preengage", "y2_prevalue", "y2_premotivate", "y2_preassist", "y2_preimprove", "y2_prethink", "y2_PreHealthEdNumShift", "y2_PreStudentHealthNumShift", "y2_PreTeacherRespNumShift", "y2_PreIntegrateNumShift", "y2_PreParticipateNumShift", "y2_PreImpactNumShift", "y2_PreAchievementNumShift", "y2_PrePromoteNumShift", "y2_PreEngageNumShift", "y2_PreValueNumShift", "y2_PreMotivateNumShift", "y2_PreAssistNumShift", "y2-PreImproveNumShift", "y2_PreThinkNumShift", "y2_PreWtPersonalHealthSubScore", "y2_PostWtPersonalHealthSubScore", "y2_PreWtPhysActivitySubscore", "y2_PostWtPhysActivitySubscore", "y2_PreWtNutritionSubscore", "y2_PostWtNutritionSubscore", "y2_prepd_2stress", "y2_prepd_3flu", "y2_prepd_4nutrition", "y2_prepd_5meditatio", "y2_prepd_6smoking", "y2_prepd_7ergo", "y2_prepd_8other", "y2_prepd_other", "y2_postage", "y2_postrace", "y2_postrace_other","y2_posthispanic", "y2_postgender", "y2_posthispanic", "y2_postgender", "y2_posthealth", "y2_posthealthinv", "y2_PostHealthNumShift","y2_postphyhealth", "y2_postmentalhealth", "y2_poststress", "y2_poststressinv", "y2_PostStressInvNumShift", "y2_postjobstress", "y2_postsmoke", "y2_postphyactive", "y2_postsit", "y2_postcomputer", "y2_postsoda", "y2_postfruit")
```

```{r}
drops2 <- c("y2_postwater","y2_postbreakfast","y2_postvegetables", "y2_postdeliver_nuted", "y2_postopen_nuted","y2_postother_nuted", "y2_postother_nuted2",  "y2_posthealthed", "y2_poststudenthealth", "y2_postteacherresp", "y2_postintegrate", "y2_postparticipate", "y2_postimpact", "y2_postachievement",  "y2_postpromote", "y2_postengage", "y2_postvalue", "y2_postmotivate", "y2_postassist",   "y2_postimprove", "y2_postthink", "y2_PostHealthEdNumShift", "y2_PostStudentHealthNumShift", "y2_PostTeacherRespNumShift", "y2_PostIntegrateNumShift", "y2_PostParticipateNumShift", "y2_PostImpactNumShift", "y2_PostAchievementNumShift", "y2_PostPromoteNumShift", "y2_PostEngageNumShift", "y2_PostValueNumShift",  "y2_PostMotivateNUmShift", "y2_PostAssistNumShift", "y2_PostImproveNumShift", "y2_PostThinkNumShift", "y2_PreWtHealthEdBelSelfEffSubscore", "y2_PostWtHealthEdBelSelfEffSubscore", "y2_postpd_1physical", "y2_postpd_2stress", "y2_postpd_3flu", "y2_postpd_4nutrition", "y2_postd_5meditatio", "y2_postpd_6smoking", "y2_postpd_7ergo", "y2_postpd_other", "y2_postopen_nuted2", "y2_posthealthed", "y2_poststudenthealth", "y2_postteacherresp", "y2_postintegrate", "y2_postparticipate", "y2_postimpact", "y2_postachievement", "y2_postpromote", "y2_postengage", "y2_postvalue", "y2_postmotivate", "y2_postassist", "y2_postimprove", "y2_postthink", "y2_program_pd1", "y2_program_pd2", "y2_program_pd3", "y2_program_pd4", "y2_program_kits", "y2_program_ped", "y2_program_comm", "y2_program_notchange", "y2_program_change", "y2_PostHeight_Feet")

```

```{r include=FALSE}
#drop the columns contained in the variables "drops and drops2
df <- teacher_clean[,!(names(teacher_clean) %in% drops)]
df<- df[,!(names(df) %in% drops2)]
#df will have 203 observarions and 279 variables

```

```{r include=FALSE}
#repeat same steps from above
drops3 <- c("Y1_PreHealthEdNumShift", "Y1_PreParticipateNumShift", "Y1_PostRace", "y1_PostHispanic", "Y1_PostGender", "Y1_PostStressInv", "Y1_PostVegetables", "Y1_PostOpen_NutEd2", "Y1_PreTobProduct", "Y1_PreVegetables", "Y1_PreTeacherResp", "Y1_PreIntegrate", "Y1_PreParticipate", "Y1_PreImpact", "Y1_PreAchievement", "Y1_PrePromote", "Y1_PreEnage", "Y1_PreValue", "Y1_PreMotivate", "Y1_PreAssist", "Y1_PreImprove", "Y1_PreThink", "Y1_PreHealthEdNumShift", "Y1_PreParticipateNumShift", "Y1_PreOpen_NutEd2", "Y1_PostTeacherResp", "Y1_PostIntegrate", "Y1_PostParticipate", "Y1_PostImpact", "Y1_PostAchievement", "Y1_PostPromote", "Y1_PostEnage", "Y1_PostValue", "Y1_PostMotivate", "Y1_PostAssist", "Y1_PostImprove", "Y1_PostThink", "Y1_PostHealthEdNumShift", "Y1_PostParticipateNumShift", "Y1_PrePost", "Y1_Program_Kits", "Y1_Program_Ped", "y2_preopen_nuted2", "y2_prepd_1physical", "y2_postmentalhealth", "y2_PostMotivateNumShift", "Y1_PostHeight_Inches", "Y1_PostHeight_Feet", "y2_PostHeight_Inches", "Y1_PreHeight_Feet", "Y1_PreHeight_Inches", "Y1_PostHispanic", "Y1_PostHypertension", "Y1_PostHealthEd", "Y1_PostSleep", "Y1_PostStudentHealth", "Y1_PostEngage", "Y1_PreHealthEd", "Y1_PreStudentHealth", "Y1_PostHeight_Total_Inches", "y2_preheight_feet", "y2_preheight_inches", "y2_preheight_total_inches", "y2_previg_intensity", "y2_prevegetables", "y2_PreImproveNumShift", "y2_postmentalhealth", "y2_postpd_5meditatio" )

df<- df[,!(names(df) %in% drops3)]
#now df will have 203 observarions and 219 variables 

```

```{r include=FALSE}
#repeat again
drops4 <- c("y3_prerace_other", "y3_prerace_2", "y3_prerace_3", "y3_prehispanic", "y3_prephyhealth", "y3_prehealth", "y3_prementalhealth", "y3_prestress", "y3_presmoke", "y3_prephyactive", "y3_presit", "y3_precomputer", "y3_presoda", "y3_prefruit", "y3_prevegetables", "y3_prewater", "y3_prebreakfast", "y3_predeliver_nuted", "y3_preopen_nuted", "y3_preother_nuted", "y3_preopen_nuted2", "y3_prehealthed", "y3_prestudenthealth", "y3_preteacherresp", "y3_preintegrate", "y3_preparticipate", "y3_preimpact", "y3_preachievement", "y3_prepromote", "y3_preengage", "y3_prevalue", "y3_premotivate", "y3_preassist", "y3_preimprove", "y3_prethink", "y3_prepd_1physical", "y3_prepd_2stress", "y3_prepd_3flu", "y3_prepd_4nutrition", "y3_prepd_5meditatio", "y3_prepd_6smoking", "y3_prepd_7ergo", "y3_prepd_8other", "y3_prepd_other", "ys_PreStressInv", "y3_PreStressInvNumShift", "y3_PreHealthNumShift", "y3_PreHealthEdNumShift", "y3_PreStudentHealthNumShift", "y3_PreTeacherRespNumShift", "y3_PreIntegrateNumShift", "y3_PreParticipateNumShift", "y3_PreImpactNumShift", "y3_PreAchievementNumShift", "y3_PrePromoteNumShift", "y3_PreEngageNumShift", "y3_PreValueNumShift", "y3_PreMotivateNUmShift", "y3_PreAssistNumShift", "y3_PreImproveNumShift", "y3_PreThinkNumShift")

drops5 <- c("y3_post", "y3_postage", "y3_postrace", "y3_postrace_other", "y3_postrace_2", "y3_postrace_3", "y3_posthispanic", "y3_postphyhealth", "y3_posthealth", "y3_postmentalhealth", "y3_poststress", "y3_postsmoke", "y3_postphyactive", "y3_postsit", "y3_postcomputer", "y3_postsoda", "y3_postfruit", "y3_postvegetables", "y3_postwater", "y3_postbreakfast", "y3_postdeliver_nuted", "y3_postopen_nuted", "y3_postother_nuted", "y3_postopen_nuted2", "y3_posthealthed", "y3_poststudenthealth", "y3_postteacherresp", "y3_postintegrate", "y3_postparticipate", "y3_postimpact", "y3_postachievement", "y3_postpromote", "y3_postengage", "y3_postvalue", "y3_postmotivate", "y3_postassist", "y3_postimprove", "y3_postthink", "y3_postpd_1physical", "y3_postpd_2stress", "y3_postpd_3flu", "y3_postpd_4nutrition", "y3_postpd_5meditatio", "y3_postpd_6smoking", "y3_postpd_7ergo", "y3_postpd_8other", "y3_postpd_other", "ys_PostStressInv", "y3_PostStressInvNumShift", "y3_PostHealthNumShift", "y3_PostHealthEdNumShift", "y3_PostStudentHealthNumShift", "y3_PostTeacherRespNumShift", "y3_PostIntegrateNumShift", "y3_PostParticipateNumShift", "y3_PostImpactNumShift", "y3_PostAchievementNumShift", "y3_PostPromoteNumShift", "y3_PostEngageNumShift", "y3_PostValueNumShift", "y3_PostMotivateNUmShift", "y3_PostAssistNumShift", "y3_PostImproveNumShift", "y3_PostThinkNumShift", "y3_PreMotivateNumShift", "y3_postmentalhealth", "y3_postgender", "y3_postjobstress", "y3_posthealtheducationbeliefsmean", "y3_program_notchange", "y3_program_change", "y3_program_comm", "y3_program_ped", "y3_program_kits", "y3_program_pd4", "y3_program_pd3", "y3_program_pd2", "y3_program_pd1", "y3_prepost", "y3_postselfefficacymean", "y3_kit", "y3_lessonstaught", "y3_prepost", "y3_postmentalhealth")
```

```{r include=FALSE}
df<- df[,!(names(df) %in% drops4)]
#now will have 203 observations and 161 variables 


#Last time
teacher_final<- df[,!(names(df) %in% drops5)]
##view(teacher_final)
# A filter Column was created when we dropped variables. We will now get rid of that column 
teacher_final<- teacher_final %>%
  dplyr::select(-filter_.)

#this teacher_final dataset will have 203 observations and 97 variables
rm(df)
```


##Section 3: Standardize Varianle Names. 
```{r include=FALSE}
#duplicate the "Code" variable and School_Type variable because we are going to move into splitting up the teacher data into 2 datasets, 1 for year 1 and 1 for year 2 data

teacher_final$y2_Code = teacher_final$Code
teacher_final$y3_Code = teacher_final$Code
teacher_final$y2_School_Type = teacher_final$School_Type
teacher_final$y3_School_Type = teacher_final$School_Type

```

##Section 4: Split Teacher Data
```{r include=FALSE}
#now any variable with a y2 infront of it we will split into a seperate data set
teacher_y2 <- teacher_final %>% 
  dplyr::select(contains("y2")) #this dataset should have 203 observations and 26 variables

teacher_y3 <- teacher_final %>%
  dplyr::select(contains("y3")) #203 observations and 23 variables

#remove y2s from the teacher_clean data set to create a y1 dataset 
teacher_y1 <- teacher_final %>% dplyr::select(!contains("y2"))
teacher_y1 <- teacher_y1%>% dplyr::select(!contains("y3")) #this dataset will have 203 observations and 52 variables 
colnames(teacher_y1)
```
###Step 5: Filter some Teachers
```{r include=FALSE}
#teacher_y1 <- teacher_y1 %>%
  #filter(Y1_Pre == "Yes" | Y1_Post == "Yes")

#teacher_y2 <- teacher_y2 %>%
  #filter(y2_pre == "1" | y2_post == "1")
#teacher_y3<- teacher_y3 %>%
  #filter(y3_pre == "1")

#before merging datasets, write individual csvs of the created teacher datasets by year (because they only recieved pdsessions in year 1) make for year 2,3, 4, 5
teacher_y1$Y1_PDSessions[is.na(teacher_y1$Y1_PDSessions)==TRUE] <- 0
teacher_y1$Y1_LessonsTaught[is.na(teacher_y1$Y1_LessonsTaught)==TRUE] <- 0
#write.csv(teacher_y1, "teacher_y1.csv")
#write.csv(teacher_y2, "teacher_y2.csv")
#write.csv(teacher_y3, "teacher_y3.csv")

```

```{r include=FALSE}
rm(teacher_clean, teacher_final)
```

##Section 5: Create Datasets to Merge
```{r include=FALSE}
teacher_y1_melt<- teacher_y1
teacher_y2_melt<- teacher_y2
teacher_y3_melt<- teacher_y3
#next is year 4 and 5
```

###Step 7:Standardize variable names by removing Y1_ or Y2_ or Y3 from colnames to prep for merge.  
```{r include=FALSE}

# Make Sure that Teacher Data Variables are the same 
#colnames(teacher_y1_melt)
#colnames(teacher_y2_melt)
colnames(teacher_y1_melt)<- gsub("Y1_", "", colnames(teacher_y1_melt))
colnames(teacher_y1_melt)<- gsub("y1_", "", colnames(teacher_y1_melt))


colnames(teacher_y2_melt)<- gsub("Y2_", "", colnames(teacher_y2_melt))
colnames(teacher_y2_melt)<- gsub("y2_", "", colnames(teacher_y2_melt))

colnames(teacher_y3_melt)<- gsub("Y3_", "", colnames(teacher_y3_melt))
colnames(teacher_y3_melt)<- gsub("y3_", "", colnames(teacher_y3_melt))
```

###Step 8: Create Program Year 
```{r include=FALSE}
#create new variable called program year 
teacher_y3_melt$Program_Year <- 3 
teacher_y2_melt$Program_Year<- 2
teacher_y1_melt$Program_Year<- 1

```
###Step 9: Column names to Lower then Check Column Names Across Datasets. 
```{r include=FALSE}

# Columns to lower for merge
colnames(teacher_y1_melt) <- tolower(colnames(teacher_y1_melt))
colnames(teacher_y2_melt) <- tolower(colnames(teacher_y2_melt))
colnames(teacher_y3_melt) <- tolower(colnames(teacher_y3_melt))

#Now determine what variables exist in one dataset but not the others

names1<- colnames(teacher_y1_melt)
names2<- colnames(teacher_y2_melt)
names3<- colnames(teacher_y3_melt)

names1[(names1 %in% names2)==FALSE]
names2[(names2 %in% names1)==FALSE] #delete post and prepost, but rename prehealtheducationbeliefmean
names1[(names1 %in% names3)==FALSE]
names3[(names3 %in% names1) == FALSE] #Rename prestressinv to prestress, rename prehealtheducationbeliefmean

```

```{r include=FALSE}
#remove the columns (after you run this, go back and rerun the definig names and check to see if actually removed)

teacher_y3_melt$prestress<- teacher_y3_melt$prestressinv
teacher_y3_melt<- teacher_y3_melt %>%
  dplyr::select(-prestressinv,-pre)

#remove the columns
teacher_y2_melt <- teacher_y2_melt %>%
  dplyr::select(-c(post, prepost, pre))

#rename this variable in the y1 dataset because the other ones are correct
teacher_y1_melt$prehealtheducationbeliefmean <- teacher_y1_melt$prehealtheducationbeliefsmean

#remove the columns
teacher_y1_melt <- teacher_y1_melt %>%
  dplyr::select(-prehealtheducationbeliefsmean)

# teacher_y1_melt has both prehealtheducationbeliefs and prehealtheducationbelief - They Are The Same - so we remove Beliefs

teacher_y2_melt<- teacher_y2_melt %>%
  dplyr::select(-prehealtheducationbeliefsmean)
```


##Section 6: Create TeacherID 
```{r include=FALSE}
teacher_y1_melt$TeacherID<- as.factor(teacher_y1_melt$code)

teacher_y1_melt<- teacher_y1_melt %>%
  dplyr::select(-code)

str(teacher_y2_melt$code)
teacher_y2_melt$TeacherID<- as.factor(teacher_y2_melt$code)

teacher_y2_melt<- teacher_y2_melt %>%
  dplyr::select(-code)

teacher_y3_melt$TeacherID<- as.factor(teacher_y3_melt$code)
teacher_y3_melt<- teacher_y3_melt %>%
  dplyr::select(-code)
#The data is now ready to combine to create Teacher_Year variables.  
```

###Step 10: Full Join
```{r include=FALSE}
#do a full join on the datasets
y1_y2 <- full_join(teacher_y1_melt, teacher_y2_melt)
#Prestress is as character in y1_Y2 so fix
y1_y2$prestress<- as.integer(y1_y2$prestress)
teacher_all <- full_join(y1_y2, teacher_y3_melt)
teacher_all %>%
  filter(is.na(school)==TRUE)
#filter out schools that don't have a name attached - these are blank across all variables except teacher ID 
#view(teacher_all)
teacher_all<-teacher_all %>%
filter(school != "NA" | lessonstaught !="NA"|pdsessions >0) 


#should have 281 observations and 53 variables
#Check Different Variables to make sure we didn't lose any important data 

#Read in first: table(teacher_data$y3_PDSessions)
#table(teacher_all$pdsessions,teacher_all$program_year)
```
##Section 7: Tidying Up
```{r include=FALSE}
#Some schools are missing the type so need to recode them
teacher_all$school_type[teacher_all$school== "Moten"] <- "Experimental"
teacher_all$school_type[teacher_all$school== "JC Nalle"] <- "Experimental"
teacher_all$school_type[teacher_all$school== "Houston"] <- "Control"
teacher_all$school_type[teacher_all$school== "Turner"] <- "Control"
#ask for school names for year 4 and 5 

#need to recode the pdsessions and lessons taught to equal zero instead of NA
teacher_all$pdsessions[is.na(teacher_all$pdsessions) == TRUE] <- 0
teacher_all$lessonstaught[is.na(teacher_all$lessonstaught)==TRUE] <- 0

#view(teacher_all) 
#should have 281 obs and 53 variables.

```

```{r include=FALSE}
#reove datasets 
rm(y1_y2)
```

```{r include=FALSE}
# Write Teacher_all 
write.csv(teacher_all, "teacher_all.csv")

```

```{r echo=FALSE}
head(teacher_all)
```



